---
title: "ST443 Lab02/pc"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Linear regression

We follow Section 3.6 of ISLR

- Simple linear regression (SLR)
- multiple linear regression (MLR)
- inference for regression
- diagnostic checking

We will analyse some data in the MASS and ISLR packages. 

```{r cars}
# install.packages("MASS")
# install.packages("ISLR")

library(MASS)
library(ISLR)
```

## 1. Simple Linear Regression

Load the Boston Housing data (that is in MASS), and

- look at the documentation of the data
- look at the list of variables in the data
- regress _medv_ on _lstat_ and interpret the summary results
- output the confidence intervals of the coefficients and interpret
- output the confidence interval and prediction interval for $E(medv)$ given a new input 

```{r}
reg <- lm(medv ~ lstat, data = Boston)

confint(reg, "(Intercept)", level = 0.95)
confint(reg, "lstat", level = 0.95)

new_data <- data.frame(lstat = 8:15)
predict.lm(reg, new_data)
```

Predictive values, confidence intervals and prediction intervals for the prediction of _medv_ for a given value of _lstat_
$CI(\hat{y}_i)  = CI(\hat{\beta}_0 + x_i \cdot \hat{\beta}_1)$,
$PI(\hat{y}_i)  = CI(\hat{\beta}_0 + x_i \cdot \hat{\beta}_1 + e_i)$,

```{r}

```

**Activity 1**
Repeat the demo by regressing _medv_ against another continuous predictor.

```{r}
#Code

reg <- lm(medv ~ tax, data = Boston)

confint(reg, "tax", level = 0.95)
confint(reg, "(Intercept)", level = 0.95)
```



## 1.2 Diagnostic checks
Under the standard assumptions, OLS estimator is BLUE, and with the normality assumption included OLS is BUE, and the t and F tests are exact.

Assumptions: "NICE": normality, independent, constant variance, $E{(e)} = 0$ + no influential points


The basic plot is the residual plot - it plots the (standardized) residuals vs fitted values or on the predictor(s).
```{r}

```

Any of the following in the residual plot suggests the model can be improved:
-(regression) outliers as these may be influential points
- some pattern. This means omission of some relevant Xs, function form for $E(Y|x)$ is not appropriate
- residuals do not fall in a band around zero. This means the assumption of constant variance of error term is not met. We call this heteroscasticity.

Is the error term normally distributed? We can get some idea using the normal QQ-plot or histogram

```{r}

```

The function `plot.lm' gathers together 6 diagnostic plots. Not all are popularly used.

```{r}

```


**Activity 2**
For the model in Activity 1, examine the diagnostic plots
```{r}
#code
```



## 3.  Multiple Linear Regression
In this demo:
- fit a MLR on a set of Xs
- fit a MLR on all the Xs (short-cut)
- compare two nested models using the F-test
- add an interaction term

Let's say we want to regress _medv_ on _lstat_ and _age_. 

```{r}

```

Say we wish to include the interaction of _ptratio_ and _tax_.

```{r}

```

**Activity 3**
1. Fit a MLR of _medv_ on some of the predictors. Think about what are the sensible predictors to include. 
2. Think also what 2 way interaction term may be present, and include this in the model. Interpret the interaction term. Is it statistically significant?
```{r}
#code

```


## 4. Variables Transformation

Is there a diminishing return on price with the number of bedrooms? We can see if there is such a phenomenon by including a squared term of `rm'.

`I()` is required to isolate the transformation on `rm'. `I()` inhibits the interpretation of operators such as "+", "-", "*" and "^" as formula operators when the appear to the right of ~ in `lm', so they are used as arithmetical operators.

```{r}

```

**Activity 4**
1. What is the min and max number of bedrooms for a property in the data?
2. Conduct a statistical test to determine whether there is support of a dimishing effect on _medv_ with `rm'.`

```{r}
#code

```


## 5 Categorical Predictors

Categorical predictors, known as factors in R, may be included in regression. A regression model can contain a mix of continuous and factors.

The Carseat data in ISLR is simulated data set containing sales of child car seats at 400 different stores.

```{r}
?Carseats
str(Carseats)
```

There are 3 factors. By default factors go into a regression as a set of levels, where one level is treated as the reference/baseline level. By default the reference level is the one that comes first when the levels are arranged in alphabetical order.

Can use `constrasts' to check.

```{r}


```
Let's run a MLR including the factor `Urban'.

```{r}

```

**Activity 5**
1. Fit a MLR for _sales_ including one or more factors. Look at the sign of the estimated coefficient on the level(s) of the factor. Is this the sign you woud expect?
2. Include a 2 way interaction where the Xs are both factors. Try it with `Urban' and `US' as both these have 2 levels so will be eaier to interpret. 


## 6. Writing Functions
```{r}
f1 = function(x){
  d = median(x) - mean(x)
  d
  #return(d)
}
f1(x = c(1,2,6))
```

```{r}
# LoadLibraries
# LoadLibraries()
LoadLibraries = function() {
  library(ISLR)
  library(MASS)
  print("The libraries have been loaded.")
}
LoadLibraries
LoadLibraries()
```